---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at <https://ds202-at-isu.github.io/labs.html> for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.

All submissions to the github repo will be automatically uploaded for grading once the due date is passed. Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
library(tidyr)
library(dplyr)
library(tidyverse)

# Assuming your dataset is named 'av'
# Replace 'path_to_data' with the actual path to your data file if it's not loaded in R yet

# Load data
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
  # Replace with your actual data file


# # Reshape Death columns
# deaths <- av %>%
#   pivot_longer(cols = starts_with("Death"),
#                names_to = c(".value", "Time"),
#                names_pattern = "([A-Za-z]+)([0-9]+)") %>%
#   mutate(Death = ifelse(Death != "", "yes", "no"))  # Mark as "yes" if not empty, otherwise "no"
# 
# # Reshape Return columns
# returns <- av %>%
#   pivot_longer(cols = starts_with("Return"),
#                names_to = c(".value", "Time"),
#                names_pattern = "([A-Za-z]+)([0-9]+)") %>%
#   mutate(Return = ifelse(Return != "", "yes", "no"))  # Mark as "yes" if not empty, otherwise "no"
# 
# # Calculate average deaths and returns suffered by an Avenger
# average_deaths <- deaths %>%
#   group_by(Name.Alias) %>%
#   summarise(avg_deaths = mean(Death == "yes", na.rm = TRUE))
# 
# average_returns <- returns %>%
#   group_by(Name.Alias) %>%
#   summarise(avg_returns = mean(Return == "yes", na.rm = TRUE))
# 
# # Output the results
# print("Average deaths:")
# print(average_deaths, n = nrow(average_deaths))
# print("Average returns:")
# print(average_returns, n = nrow(average_returns))
```

```{r}

library(tidyr)
library(dplyr)
library(tidyverse)


deaths <- av %>% select(
  URL,
  Name.Alias,
  starts_with("Death")
) %>% pivot_longer(
  cols = Death1:Death5,
  names_to = "Time",
  values_to = "Died"
)
deaths
deaths <- deaths %>% mutate(
  Time = parse_number(Time)
)
maxdeaths <- deaths %>%
  filter(Died != "") %>%
  group_by(URL, Died) %>%
  summarise(
    Time_max = max(Time, na.rm = TRUE),
    .groups = "drop"
  )

maxdeaths %>% count(Time_max, Died)
deathsavg <-  maxdeaths %>% filter(Died == "YES")
deathsavg %>% count(Time_max, Died)

returns <- av %>% select(
  URL,
  Name.Alias,
  starts_with("Return")
) %>% pivot_longer(
  cols = Return1:Return5,
  names_to = "Time",
  values_to = "Return"
)
returns
returns <- returns %>% mutate(
  Time = parse_number(Time)
)
maxreturns <- returns %>%
  filter(Return != "") %>%
  group_by(URL, Return) %>%
  summarise(
    Time_max = max(Time, na.rm = TRUE),
    .groups = "drop"
  )

maxreturns %>% count(Time_max, Return)

deathsavg %>% count(Time_max, Died) -> tt
tt
# A tibble: 4 × 3

tt$Time_max
tt$n
tt$Time_max * tt$n

sum(tt$Time_max * tt$n)
numdeath <- sum(tt$Time_max * tt$n)
print(numdeath/173)


jocastadeaths <- deaths %>%
  filter(Died != "") %>%
  filter(Name.Alias == "Jocasta") %>%
  group_by(URL, Died) %>%
  summarise(
    Time_max = max(Time, na.rm = TRUE),
    .groups = "drop"
  )
jocastareturns <- returns %>%
  filter(Return != "") %>%
  filter(Name.Alias == "Jocasta") %>%
  group_by(URL, Return) %>%
  summarise(
    Time_max = max(Time, na.rm = TRUE),
    .groups = "drop"
  )

jocastadeaths %>% count(Time_max, Died)
jocastareturns %>% count(Time_max, Return)


numdeath <- sum(tt$Time_max * tt$n)
print(numdeath/173)

```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`.

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers.

**There is an average of .51 deaths per Avenger.**

## Individually

For each team member, copy this part of the report.

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

```{r}
#FiveThirtyEight Statement: Gabriel Wright
#"and on 57 occasions the individual made a comeback."

#Code of individuals returned after their first death
totalreturns <- maxreturns%>% filter(Return == "YES")
totalreturns %>% count(Time_max, Return) -> rr


deathsavg %>% count(Time_max, Died) -> tt
rr$Time_max
rr$n
rr$Time_max * rr$n
sum(rr$Time_max * rr$n)
#acccording to our code the analysis is correct that there are 57 total returns
```

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.

## Cassie

For each team member, copy this part of the report.

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> "There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife."

### Include the code

```{r}
maxreturns %>% count(Time_max, Return) 
maxreturns %>% count(Time_max, Return) -> rr

deathsavg %>% count(Time_max, Died) -> cc
cc
# A tibble: 4 × 3

cc$Time_max
cc$n
cc$Time_max * cc$n

sum(cc$Time_max * cc$n)
numdeath <- sum(cc$Time_max * cc$n)


```

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

This shows that out of the 53 avengers that died once, 38 of them came back which is 71.69% chance they cam back from the dead the first time. This is slightly higher then what the article states.

Upload your changes to the repository. Discuss and refine answers as a team.
